---
title: 'Programmieren mit R: Seminararbeit 3'
author:
- Daniyar Akhmetov (5127348)
- Marcelo Rainho Avila (4679876)
- Xuan Son Le (4669361)
date: 'Abgabedatum: 08.02.2018'
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
graphics: yes
lof: no
fontsize: 12
documentclass: article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Part I: *Linear regression* (15 points)
  
## *Raw implementation*
```{r}
linModEst <- function(x,y) {
  ## compute temp=(x'x)^(-1) with x'=t(x)
  temp <- crossprod(x,y = NULL) #compute (x'x)
  temp <- solve(temp) #compute (x'x)^(-1)
  
  ## compute beta
  temp1 <- crossprod(x,y) #compute(x'y)
  beta <- crossprod(t(temp),temp1) #compute beta
  
  ## calculate degree of freedom
  df <- nrow(x) - ncol(x)
  
  ## calculate sigma^2
  SSR <- 0
  help <- 0
  for (i in (1:nrow(x))) {
    help <- (x[i,]) %*% beta
    SSR <- SSR + (y[i] - help)^2
  }
  sigma_power_2 <- as.double(SSR / df)
  sigma <- sqrt(sigma_power_2)
  
  # calculate covariance matrix
  vcov <- sigma_power_2 * temp
  
  # return results
  list(beta = beta, vcov = vcov, sigma = sigma, df = df)
}

data(cats, package = "MASS")
linModEst(x = cbind(1, cats$Bwt,as.numeric(cats$Sex) - 1), 
          y = cats$Hwt)
```

* By adding a new column with all values of 1, we transformed the given matrix x into X.  
* `crossprod(x, y = NULL)` is equal to `t(x) %*% y`. Because `y = NULL` is taken to be the same matrix as `x`, the result will be `t(x) %*% x`. To find the inverse matrix of `x'x` we used the function `solve()`.  
* x is given as cbind bla, because the first column is used for `ßo`  

## *Check equivalent*
```{r}
Hwt <- cats$Hwt
Bwt <- cats$Bwt
Sex <- as.numeric(cats$Sex) - 1

lm_cat <- lm(Hwt ~ Bwt + Sex, data = cats)
coef(lm_cat)
vcov(lm_cat)
```

The same!!

## *Extend implementation*
```{r}
linMod <- function(formula, data){
  # still no idea :((
  lm(formula, data) # wäre schön ^^
}

linMod(Hwt ~ Bwt + Sex, data = cats)
```

# PartII: S3 for linear models






