---
title: 'Programmieren mit R: Seminararbeit 2'
author:
- Daniyar Akhmetov (5127348)
- Marcelo Rainho Avila (4679876)
- Xuan Son Le (4669361)
date: 'Abgabedatum: 19.12.2017'
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
graphics: yes
lof: no
fontsize: 12
documentclass: article
---
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# remove environmental variables
rm(list = ls())
```

# Part I: *Functions* (15 points)

## Functions I:

Define  a  function  which  given  an  atomic  vector x as  argument,
returns x after  removing  missing values

```{r function 1}
dropNa <- function(x) {
  # takes an atomic vector as an argument and returns it without missing values
  x[!is.na(x)]
}
```

```{r test function 1}
all.equal(dropNa(c(1, 2, 3, NA, 1, 2, 3)), c(1, 2, 3, 1, 2, 3))
```

## Functions II:

*Part I:* Write a function meanVarSdSe that takes a numeric vector x as argument. The function should return a named numeric vector that contains the mean, the variance, the standard deviation and the standard error of x. 
```{r function 2}
meanVarSdSe <- function(x){
  # takes an numeric vector as argument and returns a named numeric vector 
  # containing its mean, variance, standard deviation and standard error 
  c(mean = mean(x),
    var = var(x),
    sd = sd(x),
    se = sd(x) / sqrt(length(x))
  )
}

# test
x <- 1:100
meanVarSdSe(x)
``` 


\vspace{1cm} <!-- add 1 cm vertical space -->

*Part II:*  Look at the following code sequence. What result do you expect?

```{r expect, eval=FALSE}
x <- c(NA, 1:100)
meanVarSdSe(x)
```

The code returns NA values for each statistic computed, which is the output
of each function when using the default (FALSE) argument for `na.rm`.


```{r meanVarSdSe with na.rm}
meanVarSdSe <- function(x, ...){
  # computs mean, variance, standard deviation and standard error
  #
  # Args:
  #   x: a numeric vector
  #
  # Returns:
  #   mean, variance, standard deviation and standard error of input vector
  c(mean = mean(x, ...),
    var = var(x, ...),
    sd = sd(x, ...),
    se = sd(x, ...) / sqrt(length(which(!is.na(x))))
  )
}

# test
meanVarSdSe(x, na.rm = TRUE)
```


```{r meanVarSdSe 3 with dropNa}
meanVarSdSe <- function(x){
  # computs mean, variance, standard deviation and standard error while using 
  # dropNa function
  #
  # Args:
  #   x: a numeric vector
  #
  # Returns:
  #   mean, variance, standard deviation and standard error of input vector
  dropedNa <- dropNa(x)
  c(mean = mean(dropedNa),
    var = var(dropedNa),
    sd = sd(dropedNa),
    se = sd(dropedNa) / sqrt(length(dropedNa))
  )
}

# test
meanVarSdSe(c(x, NA))
```




## Functions III: 

Write an infix function %or% that behaves like the logical operator |

```{r logical operator}
"%or%" <- function(x,y) {
  # logical operator OR... TO BE DONE...
}
```












